---
title: "Introduction to R"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First steps  

Initially we will us the R console for some very simple operations. It will allow you to familiarize yourself with the software and quickly make you a confident user.

### Console window

For this exercise simply open up R. All you will see is the **R Console** window. Initially you can think of R as a very powerful calculator. So let us try a couple of things. Just type 4+4 and then enter

```{r }
4+4

```

I hope that you got the same result as above. Let us try another task:

```{r }
4*4

```

Okay, now you see that R can calculate numbers, but let us look at some functions. E.g. let us take the square root of 9.

```{r }
sqrt(9)

```

This is already a small example of running a function (in this case sqrt). Functions will become important as we go along. 
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
